Data Encryption Standard Implementation 4003-482-01 RIT

### 2010-05-06 paper/paper.tex
Added profiling to Original Runtime

\section{Original Runtime Measurements}
\paragraph{}Each test consisted of 1,000,000 encryptions of the message 123456789ABCDEF and key 133457799BBCDFF1. They were performed on a 1.8 Ghz Core2Duo computer running Ubuntu 10.04.
\\Test 1:
\\ \indent 10.899 seconds
\\Test 2:
\\ \indent 10.867 seconds
\\Test 3:
\\ \indent 10.870 seconds
\\\\Average:
\\ \indent 10.879 seconds
\section{Original Runtime Analysis}
\indent With a total run time of 10.884 seconds, this means each encryption took only .01 milliseconds.  To find improvements, we look at our profiling.  The most time was spent in our mixer function.  To reduce this we decided to 'unravel' the loop.  To do this we would have had to make a unique method for each of the different mixing procedures.  However instead we decided to get rid of it all together by putting the unraveled contents wherever the code was called from.
\\ \indent The next worst function was the f (feistel) function.  One of the first things we noticed could be done for improvment was to remove the first for-loop. First it mixes the input, then runs through it again doing xor with the key. We combined the two steps into the already unraveled mix procedure. Also, instead of using our get method to retrieve bit values, we copied the code into the function and hard coded the values that were needed for each use.
\\ \indent After this we analyzed the keyschedule algorithm, as this was #3 in terms of time spent executing.  We had already internalized the mixer function.  The innermost for loop was unraveld to remove time spent computing the the counter and comparisons.  Also, we turned the for-loop outside of it into a do-while loop.  This made modest savings since in cases where it only loops once, since there are no extra jumps.  At the end of the function was one of the unraveled mix algorithms.  We combined this with the for loop afterwards so that we weren't running through each element twice.
\section{Revised Software design}
\section{Revised Runtime Measurements}
\paragraph{}Each test consisted of 1,000,000 encryptions of the message 123456789ABCDEF and key 133457799BBCDFF1. They were performed on a 1.8 Ghz Core2Duo computer running Ubuntu 10.04.
\\Test 1:
\\ \indent 5.960 seconds
\\Test 2:
\\ \indent 6.012 seconds
\\Test 3:
\\ \indent 5.903 seconds
\\\\Average:
\\ \indent 5.958 seconds
\section{Revised Runtime Analysis}
\section{Developers manual}
\section{Users Manual}

>>>>

\section{Original Runtime Measurements}
\paragraph{}Each test consisted of 1,000,000 encryptions of the message 123456789ABCDEF and key 133457799BBCDFF1. They were performed on a 1.8 Ghz Core2Duo computer running Ubuntu 10.04.
\\Test 1:
\\ \indent 10.899 seconds
\\Test 2:
\\ \indent 10.867 seconds
\\Test 3:
\\ \indent 10.870 seconds
\\\\Average:
\\ \indent 10.879 seconds
\begin{verbatim}
  % of   # of      # of                
  time  Seconds    Calls    Name    
 40.93   18.00   51000000   DES::mixer()
 22.88   10.06   16000000   DES::f()
 20.17    8.87    1000000   DES::keyschedule()
  8.20    3.60    1000000   DES::algorithm()
  5.59    2.46  512000000   DES::get()
  0.61    0.27   32000000   std::swap<T>()
  0.53    0.23    1000000   std::swap_ranges<T,G>()
  0.41    0.18   32000000   std::__iter_swap<T>::iter_swap<G,E>()
  0.28    0.12   32000000   std::iter_swap<T,G>()
  0.20    0.09              DES::on()
  0.11    0.05    1000000   DES::DES()
  0.02    0.01    1000000   DES::encrypt()
  0.02    0.01    1000000   DES::~DES()
  0.02    0.01              DES::main
\end{verbatim}
\section{Original Runtime Analysis}
\indent With a total run time of 10.884 seconds, this means each encryption took only .01 milliseconds.  To find improvements, we look at our profiling.  The most time was spent in our mixer function.  To reduce this we decided to 'unravel' the loop.  To do this we would have had to make a unique method for each of the different mixing procedures.  However instead we decided to get rid of it all together by putting the unraveled contents wherever the code was called from.
\\ \indent The next worst function was the f (feistel) function.  One of the first things we noticed could be done for improvment was to remove the first for-loop. First it mixes the input, then runs through it again doing xor with the key. We combined the two steps into the already unraveled mix procedure. Also, instead of using our get method to retrieve bit values, we copied the code into the function and hard coded the values that were needed for each use.
\\ \indent After this we analyzed the keyschedule algorithm, as this was #3 in terms of time spent executing.  We had already internalized the mixer function.  The innermost for loop was unraveld to remove time spent computing the the counter and comparisons.  Also, we turned the for-loop outside of it into a do-while loop.  This made modest savings since in cases where it only loops once, since there are no extra jumps.  At the end of the function was one of the unraveled mix algorithms.  We combined this with the for loop afterwards so that we weren't running through each element twice.
\section{Revised Software design}
\section{Revised Runtime Measurements}
\paragraph{}Each test consisted of 1,000,000 encryptions of the message 123456789ABCDEF and key 133457799BBCDFF1. They were performed on a 1.8 Ghz Core2Duo computer running Ubuntu 10.04.
\\Test 1:
\\ \indent 5.960 seconds
\\Test 2:
\\ \indent 6.012 seconds
\\Test 3:
\\ \indent 5.903 seconds
\\\\Average:
\\ \indent 5.958 seconds
\section{Revised Runtime Analysis}
\section{Developers manual}
\section{Users Manual}
